Ground state energy of quantum dots using the coupled cluster method

### 2018-06-11 project_2/article/article.tex
Start on review of the theory section

    \subsection{Second quantization}
        Employing the creation operators, $\acr{p}$, and the destruction
        operators, $\ade{p}$, we can write the Hamiltonian as
        \begin{align}
            H
            &=
            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
            + \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}\acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where the sums are general indices over all $L$ basis states and the
        matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            \bra{pq}\ket{rs} \equiv \bra{pq}u\ket{rs} - \bra{pq}u\ket{sr}.
        \end{gather}
        Note that we use the ``chemists notation'' to label the antisymmetric matrix
        elements.

    \subsection{The coupled cluster approximation}
        We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined by
        \begin{align}
            \ket{\Psi_{\text{CC}}}
            &\equiv e^{T}\ket{\Phi_0}
            = \para{
                \sum_{i = 0}^n
                \frac{1}{n!}T^n
            }\ket{\Phi_0},
        \end{align}
        where the \emph{cluster operator}, $T$, is given by a sum of
        $p$-excitation operators labeled $T_p$. They consist of \emph{cluster
        amplitudes}, $t_{i\dots}^{a\dots}$, and creation and annihilation
        operators.
        \begin{align}
            T &= T_1 + T_2 + \dots + T_p \\
            &=
            \sum_{ia}t_i^a\acr{a}\ade{i}
            + \para{\frac{1}{2!}}^2\sum_{ijab}
            t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
            + \dots.
        \end{align}
        In our approximation we limit the cluster operator to only include
        double excitations,
        \begin{align}
            T \equiv T_2
            = \frac{1}{4}\sum_{ijab}t_{ij}^{ab}\acr{a}\acr{b}\ade{j}\ade{i}.
            \label{eq:T_2}
        \end{align}
        The first part of the coupled cluster method consists of constructing
        the cluster amplitudes using the \emph{amplitude equations}. After we
        have found the amplitudes we can compute the energy.

    \subsection{Energy of the coupled cluster approximation}
        When we're going to compute the energy of a system using the coupled
        cluster approximation we would ideally want to find the expectation
        value of the energy using the coupled cluster wavefunction.
        \begin{align}
            \ecc = \bcc H\kcc.
        \end{align}
        As it turns out, this is an uncomfortable way of finding the energy
        as $T \neq T^{\dagger}$. Instead we will define what we call the
        \emph{similarity transformed Hamiltonian}. We plug the coupled
        cluster wavefunction into the Schrödinger equation.
        \begin{align}
            H\kcc = \ecc\kcc.
        \end{align}
        Next, we left multiply with the inverse of the cluster expansion,
        i.e.,
        \begin{align}
            e^{-T}H\kcc = e^{-T}\ecc\kcc
            = \ecc \kslat.
            \label{eq:clean_schrodinger}
        \end{align}
        Projecting this equation on the reference state we get
        \begin{align}
            \ecc = \bslat e^{-T}H\kcc
            = \bslat e^{-T}He^{T}\kslat,
        \end{align}
        where in the latter inner-product we have located the similarity
        transformed Hamiltonian defined by
        \begin{align}
            \bar{H} \equiv e^{-T}He^{T}.
            \label{eq:similarity_transformed_hamiltonian}
        \end{align}

@@

        \subsubsection{Coupled cluster doubles energy equation}
            Using the doubles approximation with the cluster operator $T_2$
            defined in \autoref{eq:T_2} the energy equation becomes
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat e^{-T_2}H_N e^{T_2}\kslat.
            \end{align}
            As the doubles cluster operator doubly excites the reference and
            using the expansion in \autoref{eq:sim_norm_hamiltionian_expansion}
            we see that we can write the energy equation as
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat H_N \kslat + \bslat (H_N T_2)_c\kslat,
            \end{align}
            as the Hamiltonian is only able to relax one pair of single
            particle functions. By construction we have that
            \begin{align}
                \bslat H_N \kslat = 0.
            \end{align}
            In the second term only the normal ordered two-body operator can
            contribute as the cluster operator gives a total excitation of $+2$.
            As we are projecting onto the reference we have to relax to zero
            again. The normal ordered Fock operator is at most able to excite
            and relax by $1$ and does therefore not contribute to the
            overall expression.
            \begin{align}
                \bslat (W_N T_2)_c \kslat
                &= \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab}.
            \end{align}
            In total the energy equation reduces to
            \begin{align}
                \eccd
                &= \sum_{i} h_i^i + \half \sum_{ij}\bra{ij}\ket{ij}
                + \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab},
            \end{align}
            where the first two terms come from the reference energy as shown in
            \autoref{eq:reference_energy}.

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
            \label{eq:doubles_amplitude_1}
        \end{align}
        to find an expression that can be used to solve for the amplitudes 
        $t_{ij}^{ab}$.
        By employing the Baker-Campbell-Haussdorf (BCH) expansion,
        while setting $T= T_2$, we find
        \begin{equation}
                \bar{H}
                    = \Big(
                            H_N + H_NT_2
                            + \frac{1}{2}H_NT_2^2 \Big)_C.
        \label{eq:normal_order_expansion}
        \end{equation}
        The subscript $c$ indicates that only those terms in which the
        Hamiltonian is connected\footnote{Again, in aWick's theorem sense}
        to every cluster operator on its right should be included. Since
        the Hamiltonian contains at most four annihilation and creation
        operators, $H_N$ can connect to at most four cluster operators
        at once. Therefore, the BCH expansion must truncate at the
        fourth-order terms.

        Now comes the rather tedious task of evaluating all the terms
        that arises from inserting \autoref{eq:normal_order_expansion}
        into \autoref{eq:doubles_amplitude_1}. This can be done by applying
        Wick's generalised theorem, but the task is a daunting and streneous one.
        A few example computations of how this can be done is included in
        \autoref{app:wickOnAmplitude}.
         Instead of doing it in this manner, we employ the second quantisation library
         from SymPy instead\footnote{This is also more in the spirit of this project,
         as it is within the realm of \emph{Computational} Physics.}.
        The CCD amplitude equation, from SymPy computation, is
        \begin{equation}
        \begin{aligned}
            0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
                 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
                 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
                 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
                 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
        \label{eq:CCD_amp_1}
        \end{aligned}
        \end{equation}

        \subsubsection{The Iterative Scheme}
        At first, we pick only diagonal elements of $f$ to be part of the unperturbed
        Hamiltonian and consider the rest of the terms a perturbation.
        The second and third terms in \autoref{eq:CCD_amp_1} can now be
        rewritten,
        \begin{equation}
        \begin{aligned}
            f^b_c&t_{ij}^{ab}P(ab) - f^k_j t_{ik}^{ab}P(ij) \\
                &\to f^b_bt_{ij}^{ab} - f^a_at_{ij}^{ba} - f^j_jt_{ij}^{ab} + f^i_it_{ji}^{ab} \\
                &= (f^a_a + f^b_b - f^i_i - f^j_j)t_{ij}^{ab} \\
                &= (\epsilon_a + \epsilon_b - \epsilon_i - \epsilon_j)t_{ij}^{ab} \\
                &= -(\epsilon_i + \epsilon_j - \epsilon_a - \epsilon_b)t_{ij}^{ab} \\
                &= -D_{ij}^{ab}t_{ij}^{ab}.
        \end{aligned}
        \end{equation}

        Then we can define a new version of \autoref{eq:CCD_amp_1},
        \begin{equation}
        \begin{aligned}
            g(u, t) =&
                    u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
                 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
                 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
                 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
                 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
        \end{aligned}
        \end{equation}

        where $\tilde{f}$ are the non-diagonal parts of the Fock matrix.

        Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, t)$.
        This allows us to define an iterative scheme,
        \begin{equation}
            t^{(k + 1)} = \frac{g(u, t^{(k)})}{D_{ij}^{ab}},
            \label{eq:iterative_amplitude}
        \end{equation}
        with the initial guess
        \begin{equation}
            t^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
        \end{equation}

@@

        \begin{figure}
           \centering
            \includegraphics[width=244px]
            {../dat/figures/iter_run_times_zoom.pdf}
            \caption{Run time for $N=6$ particles, only the faster methods are
            included for better comparison.}
            \label{fig:running_time_6_zoom}            
        \end{figure}

@@

\section{Amplitude Equations}

    \label{app:wickOnAmplitude}

>>>>

    \subsection{Second quantization}
        Employing the creation operators, $\acr{p}$, and the destruction
        operators, $\ade{p}$, we can write the Hamiltonian as
        \begin{align}
            H
            &=
            h_{q}^{p}\acr{p}\ade{q}
            + \frac{1}{4}u^{pq}_{rs}
            \acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where we have used the Einstein summation convention with repeated
        indices having an implicit sum. The matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            u^{pq}_{rs} \equiv \bra{pq}u\ket{rs} - \bra{pq}u\ket{sr},
        \end{gather}
        where the $u$-matrix consists of the the antisymmetric two-body
        elements.

    \subsection{The coupled cluster approximation}
        We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined by
        \begin{align}
            \ket{\Psi_{\text{CC}}}
            &\equiv e^{T}\ket{\Phi_0}
            = \para{
                \sum_{i = 0}^n
                \frac{1}{n!}T^n
            }\ket{\Phi_0},
        \end{align}
        where the \emph{cluster operator}, $T$, is given by a sum of
        $p$-excitation operators labeled $T_p$. They consist of \emph{cluster
        amplitudes}, $t_{i\dots}^{a\dots}$, and creation and annihilation
        operators.
        \begin{align}
            T &= T_1 + T_2 + \dots + T_p \\
            &=
            t_i^a\acr{a}\ade{i}
            + \para{\frac{1}{2!}}^2
            t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
            + \dots.
        \end{align}
        In our approximation we limit the cluster operator to only include
        double excitations,
        \begin{align}
            T \equiv T_2
            = \frac{1}{4}t_{ij}^{ab}\acr{a}\acr{b}\ade{j}\ade{i}.
            \label{eq:T_2}
        \end{align}
        The first part of the coupled cluster method consists of constructing
        the cluster amplitudes using the \emph{amplitude equations}. After we
        have found the amplitudes we can compute the energy.

    \subsection{Energy of the coupled cluster approximation}
        When we're going to compute the energy of a system using the coupled
        cluster approximation we would ideally want to find the expectation
        value of the energy using the coupled cluster wavefunction.
        \begin{align}
            \ecc = \bcc H\kcc.
        \end{align}
        As it turns out, this is an uncomfortable way of finding the energy
        as $T \neq T^{\dagger}$. Instead we will define what we call the
        \emph{similarity transformed Hamiltonian}. We plug the coupled
        cluster wavefunction into the Schrödinger equation.
        \begin{align}
            H\kcc = \ecc\kcc.
        \end{align}
        Next, we left multiply with the inverse of the cluster expansion,
        i.e.,
        \begin{align}
            e^{-T}H\kcc = e^{-T}\ecc\kcc
            = \ecc \kslat.
            \label{eq:clean_schrodinger}
        \end{align}
        Projecting this equation onto the reference state we get
        \begin{align}
            \ecc = \bslat e^{-T}H\kcc
            = \bslat e^{-T}He^{T}\kslat,
        \end{align}
        where in the latter inner-product we have located the similarity
        transformed Hamiltonian defined by
        \begin{align}
            \bar{H} \equiv e^{-T}He^{T}.
            \label{eq:similarity_transformed_hamiltonian}
        \end{align}

@@

        \subsubsection{Coupled cluster doubles energy equation}
            Using the doubles approximation with the cluster operator $T_2$
            defined in \autoref{eq:T_2} the energy equation becomes
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat e^{-T_2}H_N e^{T_2}\kslat.
            \end{align}
            The doubles cluster operator will excite a pair of single particle
            functions in the reference state. As the single particle functions
            are orthonormal we have that
            \begin{align}
                \braket{\Phi^{X}}{\Phi_0} = 0,
            \end{align}
            where $X$ is an excitation different from zero. This means that for
            a term in the Baker-Campbell-Hausdorff expansion contribute with a
            nonzero value to the energy equation it must leave the reference
            state in its original state. The two-body part of the Hamiltonian is
            only able to relax a pair of single particle functions in the
            reference state whereas the doubles cluster operator will excite a
            pair. This means that we are only left with the following
            contributions to the energy equation.
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat H_N \kslat + \bslat (H_N T_2)_c\kslat.
            \end{align}
            Further, by construction we have that
            \begin{align}
                \bslat H_N \kslat = 0.
            \end{align}
            In the second term only the normal ordered two-body operator can
            contribute as the cluster operator gives a total excitation of $+2$.
            As we are projecting onto the reference we have to relax to zero
            again. The normal ordered Fock operator is at most able to excite
            and relax by $1$ and does therefore not contribute to the
            overall expression.
            \begin{align}
                \bslat (W_N T_2)_c \kslat
                &= \frac{1}{4}u^{ij}_{ab}t^{ab}_{ij}.
            \end{align}
            In total the energy equation reduces to
            \begin{align}
                \eccd
                &= h^i_i + \half u^{ij}_{ij}
                + \frac{1}{4} u^{ij}_{ab}t^{ab}_{ij},
                \label{eq:ccd_energy}
            \end{align}
            where the first two terms come from the reference energy. See
            Appendix \ref{app:energy_equation} for the calculation of this
            expression.

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. As in the case of the energy equation we use the
        Baker-Campbell-Hausdorff expansion when expanding the exponentials in
        the similarity transformed Hamiltonian. We only keep the terms which
        excites the reference state to the same degree as the amplitudes we are
        trying to find.

        \subsubsection{Coupled cluster doubles amplitude equations}
            In our case we are only interested in the second order ampltiudes
            found in the doubles approximation, hence we will solve the equation
            \begin{align}
                \bra{\Phi_{ij}^{ab}}e^{-T_2}H_N e^{T_2}\kslat = 0,
                \label{eq:doubles_amplitude_1}
            \end{align}
            to find an expression that can be used to solve for the amplitudes
            $t_{ij}^{ab}$. As the state we are projecting onto is doubly excited
            we will only keep terms from the Baker-Campbell-Hausdorff expansions
            which leaves the reference doubly excited.
            \begin{equation}
                \bar{H} = \para{
                    H_N + H_NT_2 + \frac{1}{2}H_NT_2^2
                }_c.
                \label{eq:normal_order_expansion}
            \end{equation}
            Higher order terms will leave the reference in a too high excitation
            state.

            Now comes the rather tedious task of evaluating all the terms that
            arises from inserting \autoref{eq:normal_order_expansion} into
            \autoref{eq:doubles_amplitude_1}. This can be done by applying
            Wick's generalised theorem, but the task is a daunting and streneous
            one. A few example computations of how this can be done is included
            in Appendix \ref{app:wick_on_amplitude}. Instead of doing it in this
            manner, we employ the second quantization library from SymPy
            \footnote{This is also more in the spirit of this project, as
            it is within the realm of \emph{Computational} Physics.}.  The CCD
            amplitude equation, from SymPy\footnote{Note that this is one of
            many ways to write the amplitude equation as the ordering of the
            indices can change drastically for every run of the SymPy-script.},
            is
            \begin{equation}
                \begin{aligned}
                    0 &= u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab)
                    - f^k_jt^{ab}_{ik}P(ij)
                    \\
                    &\qquad
                    + \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd}
                    + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd}
                    \\
                    &\qquad
                    + \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij)
                    - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab)
                    \\
                    &\qquad
                    + t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij)
                    + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij)
                    \\
                    &\qquad
                    - \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
                    \label{eq:CCD_amp_1}
                \end{aligned}
            \end{equation}
            Here $P(ij)$ is an exchange operator which interchanges two
            particles with indices $i$ and $j$.

        \subsubsection{The Iterative Scheme}
            In order to find the amplitude $t$ we have to start with an initial
            guess and use an iterative procedure to improve on the initial
            guess. We start by picking the diagonal elements of $f$ to be a part
            of the unperturbed Hamiltonian and consider the rest of the terms a
            perturbation. The second and third terms in \autoref{eq:CCD_amp_1}
            can now be rewritten,
            \begin{equation}
            \begin{aligned}
                f^b_c&t_{ij}^{ab}P(ab) - f^k_j t_{ik}^{ab}P(ij) \\
                    &\to f^b_bt_{ij}^{ab} - f^a_at_{ij}^{ba}
                    - f^j_jt_{ij}^{ab} + f^i_it_{ji}^{ab} \\
                    &= (f^a_a + f^b_b - f^i_i - f^j_j)t_{ij}^{ab} \\
                    &= (\epsilon_a + \epsilon_b - \epsilon_i
                    - \epsilon_j)t_{ij}^{ab} \\
                    &= -(\epsilon_i + \epsilon_j - \epsilon_a
                    - \epsilon_b)t_{ij}^{ab} \\
                    &= -D_{ij}^{ab}t_{ij}^{ab},
            \end{aligned}
            \end{equation}
            where the arrow signifies that we only look at the diagonal elements
            of $f$. We now define the right hand side of \autoref{eq:CCD_amp_1}
            to $g(f, u, t)$ where we make sure that we only use the off-diagonal
            elements of $f$ in the computations.

            By moving $D^{ab}_{ij}t^{ab}_{ij}$ to the left hand side we we get
            \begin{align}
                D_{ij}^{ab}t_{ij}^{ab} = g(u, t).
            \end{align}
            This allows us to define an iterative scheme,
            \begin{equation}
                t^{(k + 1)} = \frac{g(u, t^{(k)})}{D_{ij}^{ab}},
                \label{eq:iterative_amplitude}
            \end{equation}
            with the initial guess
            \begin{equation}
                t^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
            \end{equation}

@@

        \begin{figure}
           \centering
            \includegraphics[width=244px]
            {../dat/figures/iter_run_times_zoom_6.pdf}
            \caption{Run time for $N=6$ particles, only the faster methods are
            included for better comparison.}
            \label{fig:running_time_6_zoom}
        \end{figure}

@@

\section{Energy equation}
    \label{app:energy_equation}
    Here we show how to compute the coupled cluster doubles energy from
    \autoref{eq:ccd_energy}. The first two terms come from the reference energy
    shown in \autoref{eq:reference_energy}. The last term we get by using Wick's
    theorem.
    \begin{align}
        \bslat (H_N T_2)_c\kslat
        &= \bslat (W_N T_2)_c\kslat,
    \end{align}
    as the normal ordered Fock operator can at most relax a single state, i.e.,
    \begin{align}
        \bslat (F_N T_2)_c\kslat = 0.
    \end{align}
    We are thus left with computing the contraction of the doubles cluster
    operator and the normal ordered two-body operator. Note that we must keep
    only the fully contracted terms.
    \begin{align}
        \bslat (W_N T_2)_c\kslat
        &=
        \frac{1}{16} u^{pq}_{rs} t^{ab}_{ij}
        \bslat
        A
        \kslat,
        \label{eq:third_term_ccd_energy}
    \end{align}
    where $A$ is the create and destruction operator strings. By using
    generalized Wick's theorem we get
    \begin{align}
        A &=
        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
        \{\acr{a}\acr{b}\ade{j}\ade{i}\}
        \nonumber \\
        &=
        \{
            \wick{
                    \c4 a_p^{\dagger} \c3 a_q^{\dagger}
                    \c2 a_s \c1 a_r
                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
                    \c3 a_j \c4 a_i
            }
        \}
        +
        \{
            \wick{
                    \c4 a_p^{\dagger} \c3 a_q^{\dagger}
                    \c1 a_s \c2 a_r
                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
                    \c3 a_j \c4 a_i
            }
        \}
        \nonumber \\
        &\qquad
        +
        \{
            \wick{
                    \c3 a_p^{\dagger} \c4 a_q^{\dagger}
                    \c2 a_s \c1 a_r
                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
                    \c3 a_j \c4 a_i
            }
        \}
        +
        \{
            \wick{
                    \c3 a_p^{\dagger} \c4 a_q^{\dagger}
                    \c1 a_s \c2 a_r
                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
                    \c3 a_j \c4 a_i
            }
        \}
        \nonumber \\
        &=
        \delta_{pi}\delta_{qj}\delta_{sb}\delta_{ra}
        - \delta_{pi}\delta_{qj}\delta_{sa}\delta_{rb}
        \nonumber \\
        &\qquad
        - \delta_{pj}\delta_{qi}\delta_{sb}\delta_{ra}
        + \delta_{pj}\delta_{qi}\delta_{sa}\delta_{rb}.
    \end{align}
    Inserted back into \autoref{eq:third_term_ccd_energy} and summing the
    indices in the Kronecker-Delta's we get
    \begin{align}
        \bslat (W_N T_2)_c\kslat
        &=
        \frac{1}{16}\para{
            u^{ij}_{ab} - u^{ij}_{ba} - u^{ji}_{ab} + u^{ji}_{ba}
        }
        t^{ab}_{ij}
        \nonumber \\
        &=
        \frac{1}{4}u^{ij}_{ab}t^{ab}_{ij},
    \end{align}
    where we have used the antisymmetry of the two-body matrix elements, thus
    completing what we wanted to show.

\section{Amplitude equations}

    \label{app:wick_on_amplitude}
